DRAFT \\ From physical principles to classical and quantum \\ Hamiltonian and Lagrangian particle mechanics

### 2017-01-06 2015CompleteDerivation/completeDerivation.tex
Finished reviewing quantum with Christine

As we made it clear from the start, deterministic and reversible evolution is an assumption. And as we have seen in the justifications for assumptions \ref{ass:infinitesimal_reducibility} and \ref{ass:irreducibility} we cannot have determinism at all levels of decomposition while transferring information from one system to another.

@@

It is clear that it has to be absorbed as it is within the aperture of the whole array. It is also clear that it must be absorbed whole: as it is an irreducible system, the detector cannot partially interact with it, it cannot half absorb it. So a single detector will capture it. It is also clear that it cannot always be the same detector: that corresponds to states where the particle is spread within the aperture of that detector. In other words: as the particle hits the array, one detector triggers randomly based on the distribution of the particle.

The process is non-deterministic in the sense that the state of the particle is not enough to predict which detector will trigger. Maybe it depends on the state of the detector, maybe it depends on the unstated part of the particle. The point is that a single measurement does not tell us much about the state of the incoming particle.

We then repeat the process, making sure the incoming particle is in the same state. Given a single "pure" incoming particle state, we get a "mixed" statistical distribution describing the output of the array.\footnote{Note that this is different from the model used often to describe the quantum measurement problem, where the incoming system becomes entangled with the measuring device. We do not make any claim on what the final state of the particle should be: in many cases, like in a photomultiplier tube, the particle is gone and there is no state. We simply have one incoming state for the particle, prepared over and over exactly the same, and the outgoing state for the measurement device, which is a classical statistical distribution over the different measurement takes.} The question is: how is the statistical distribution coming out of the detector related to the material distribution coming in with the particle?

If the detector does not introduce biases, then the final distribution is only a function of the incoming particle: all unbiased detectors will give the same result. The unstated part also cannot affect the final distribution of outcomes: it is not physically distinguishable. The final statistical distribution of the detector must match the incoming material distribution of the particle. That is if $N$ is the total number of takes and $N_A$ is the number of times the detector with aperture in the $A$ region captured the particle, the ratio $\frac{N_A}{N}$ will be approximately the amount of material within the aperture $\int_A \psi^\dagger(q^i) \psi(q^i) d^nq$. If we assume an infinite amount of takes, we expect them to coincide: $\int_A \psi^\dagger(q^i) \psi(q^i) d^nq = \lim\limits_{N\rightarrow\infty} \frac{N_A}{N}$. But this is also the frequentist definition of probability. Therefore we have $\int_A \psi^\dagger(q^i) \psi(q^i) d^nq = P(\textrm{measuring particle in } A)$ thus recovering the probabilistic interpretation of the wave function.

We can conceptually extend this picture to the generic case. Suppose we know the outcomes of our measurement are parametrized by a fragment variable $\hat{q}$. We will have a linear functional $\Lambda^{\frac{1}{2}}_{\hat{q}=\hat{q}_0} : \Psi \rightarrow \mathbb{C}$ such that $|\Lambda^{\frac{1}{2}}_{\hat{q}=\hat{q}_0}|^2$ returns the amount of material prepared with the particular value $\hat{q}_0$. We can re-express this quantity in terms of the dual and the inner product extended to all Lebesgue integrable functions: $|\Lambda^{\frac{1}{2}}_{\hat{q}=\hat{q}_0}|^2 = |\langle \hat{q}_0 | \psi \rangle|^2 $. This is also the probability that the outcome associated with $\hat{q}_0$ is selected by the measuring device, thus recovering the statistical nature of the quantum predictions.

Note that while this process is of a non-deterministic nature, it is of a specific non-deterministic nature. It requires that no other state can influence the final distribution or the distribution of measurement outcomes would not match the distribution of incoming material. That is: it cannot be understood as an averaging over some unknown external state.

Also note that while we have a way to predict the distribution over final outcomes from the initial state, we have no way of describing what happens in between. When we assume that only the initial state matters, we are saying that we can disregard the particulars of the non-deterministic evolution. So much so that we can substitute it with a deterministic one to compute the probability distribution. In particular, we have no way of predicting which set of final outcomes we will be projecting to, nor when does that actually happen.

While this may seem like an incomplete description, it is as complete as the irreducibility assumption allows. The state is known and determines the final statistical distribution. The unstated part cannot be known, and influences the non-deterministic evolution. If we were able to say something more about the non-deterministic evolution, we would be able to describe an aspect of the unstated part, which we cannot do without violating the assumption. If we cannot learn anything about the unstated part from the final distribution, then we are consistent with our assumption and we are able to use deterministic evolution and projection to determine the final mixed state. In other words: the inability to describe the unstated part must be matched with a corresponding inability to describe its evolution.

We therefore have arrived to all the major underpinning of quantum mechanics and see how they stem from assumption \ref{ass:irreducibility} of irreducibility. This gives us a more satisfactory account on what quantum mechanics describes, though it does not claim to solve many of the open problems. It does, though, give good reason why we have those problems and how the desire to know ``what is really going on" is hindered by the very assumption that the quantum description is based on.

@@

As we went through many details in the derivation, it is time to step back and conclude with a few general remark.

The first remarkable aspect of this work is that something along these lines can be done. Regardless of whether all the minor details are addressed in the best possible way, the idea that so much can be derived from so little was not something we expected. In hindsight, the idea that physics should be founded on physical principles is nothing new: it just hasn't been thoroughly applied to basic mechanics. If this work would have been done incrementally from the start, these concepts would have been developed gradually and the results would feel more mundane.

The second striking aspect is a shift in perspective that this work seems to suggest. The usual reductionist view is that you start with the rules of the small parts and then build up the composite whole. This works shift that perspective upside down: it's not what the parts can tell you about the whole, it's what the whole can tell you about the parts.

The state space for classical particles is the one that allows to express invariant distributions: that is a requirement that comes from the whole. The fragments of quantum particles do not even have a well defined description: the whole tells us nothing about them. In retrospect, this is not be surprising. Nobody is going to hand us down the laws of infinitesimal parts: we induce them by manipulating finite systems. Therefore what the whole allows to infer about its part is physically fundamental.

@@

In this section we include mathematical demonstration in support of the physical arguments of the paper.

@@

\begin{prop}\label{symplectomorphism_generator}
	Let $(M, \omega)$ a symplectic manifold. Let $f: (M, \omega) \rightarrow (M, \omega)$ an infinitesimal self-symplectomorphism. Let $S \in TM \; | \; f(\xi^a(m)) = \xi^a(m) + S^a(m)dt \; \forall m \in M$ be the infinitesimal displacement. There exists a function $H: M \rightarrow \mathbb{R}$ such that $S^{a} \omega_{ab} = \partial_{b}H$ and $H \in C^2(M, \mathbb{R})$.
\end{prop}

>>>>

As we made it clear from the start, deterministic and reversible evolution is an assumption. And as we have seen in the rationale for assumptions \ref{ass:infinitesimal_reducibility} and \ref{ass:irreducibility} we cannot have determinism at all levels of decomposition while transferring information from one system to another.

@@

It is clear that it has to be absorbed as it is within the aperture of the whole array. It is also clear that it must be absorbed whole: as it is an irreducible system, the detector cannot partially interact with it, it cannot half absorb it. So a single detector will capture it. It is also clear that it cannot always be the same detector: that corresponds to states where the particle is spread within the aperture of that detector. In other words: as the particle hits the array, one detector triggers non-deterministically based on the distribution of the particle.

The process is non-deterministic in the sense that the state of the particle is not enough to predict which detector will trigger. Maybe it depends on the state or unstated part of the detector, maybe it depends on the unstated part of the particle or maybe on the environment. The point is that a single measurement does not tell us much about the state of the incoming particle.

We then repeat the process, making sure the incoming particle is prepared in the same way. Given a single "pure" incoming particle state, we get a "mixed" statistical distribution describing the output of the array.\footnote{Note that this is different from the model used often to describe the quantum measurement problem, where the incoming system becomes entangled with the measuring device. We do not make any claim on what the final state of the particle should be: in many cases, like in a photomultiplier tube, the particle is gone and there is no state. We simply have one incoming state for the particle, prepared over and over exactly the same, and the outgoing state for the measurement device, which is a classical statistical distribution over the different measurement takes.} The question is: how is the statistical distribution coming out of the detector related to the material distribution coming in with the particle?

If the detector does not introduce biases, then the final distribution is only a function of the incoming particle: all unbiased detectors will give the same result. The unstated part also cannot affect the final distribution of outcomes: it is not physically distinguishable. The final statistical distribution of the detector must match the incoming material distribution of the particle. That is if $N$ is the total number of takes and $N_A$ is the number of times the detector with aperture in the $A$ region captured the particle, the ratio $\frac{N_A}{N}$ will be approximately the amount of material within the aperture $\int_A \psi^\dagger(q^i) \psi(q^i) d^nq$. If we assume an infinite number of takes, we expect them to coincide: $\int_A \psi^\dagger(q^i) \psi(q^i) d^nq = \lim\limits_{N\rightarrow\infty} \frac{N_A}{N}$. But this is also the frequentist definition of probability. Therefore we have $\int_A \psi^\dagger(q^i) \psi(q^i) d^nq = P(\textrm{measuring particle in } A)$ thus recovering the probabilistic interpretation of the wave function.

We can conceptually extend this picture to the generic case. Suppose we know the outcomes of our measurement are parametrized by a fragment variable $\hat{q}$. We will have a linear functional $\Lambda^{\frac{1}{2}}_{\hat{q}=\hat{q}_0} : \Psi \rightarrow \mathbb{C}$ such that $|\Lambda^{\frac{1}{2}}_{\hat{q}=\hat{q}_0}|^2$ returns the amount of material prepared with the particular value $\hat{q}_0$. We can re-express this quantity in terms of the dual vector $\hat{q}_0$ corresponding to the linear functional $\Lambda^{\frac{1}{2}}_{\hat{q}=\hat{q}_0}$ and the inner product extended to all Lebesgue integrable functions: $|\Lambda^{\frac{1}{2}}_{\hat{q}=\hat{q}_0}|^2 = |\langle \hat{q}_0 | \psi \rangle|^2 $. This is also the probability that the outcome associated with $\hat{q}_0$ is selected by the measuring device, thus recovering the statistical nature of the quantum predictions.

Note that while this process is of a non-deterministic nature, it is of a specific non-deterministic nature. It requires that no other state can influence the final distribution as the match between the physical material distribution and the statistical measurement distribution was recovered only under the assumption that the detector does not introduce any bias. That is: the non-deterministic nature cannot be understood as an averaging over some unknown external state.

Also note that while we have a way to predict the distribution over final outcomes from the initial state, we have no way of describing what happens in between. When we assume that only the initial state matters, we are saying that we can disregard the particulars of the non-deterministic evolution. So much so that we can substitute it with a deterministic one to compute the probability distribution. In particular, given only the initial state, we have no way of predicting which set of final outcomes the system will project to as that depends on the actual choice of detector setup, nor when that projection actually happens.

While this may seem like an incomplete description, it is as complete as the irreducibility assumption allows. The state can be known and determines any final statistical distribution. The unstated part cannot be known, and is what gives the statistical distribution in the first place. If we were able to say something more about the evolution of the unstated part, we would be able to further describe it, which we cannot do without violating the assumption. If we cannot learn anything about the unstated part from the final distribution, then we are consistent with our assumption and we are able to use deterministic evolution and projection to determine the final statistical distribution. In other words: the inability to describe the unstated part not only constrains the state space and deterministic evolution, but also the set of non-deterministic processes by excluding those for which the final statistical distribution would depend on the unstated part itself.

We therefore have arrived at all the major underpinnings of quantum mechanics and see how they stem from assumption \ref{ass:irreducibility} of irreducibility. This gives us a more satisfactory account of what quantum mechanics describes, although it does not claim to solve many of the open problems. It does, however, give good reasons as to why we have those problems and how the desire to know ``what is really going on" is hindered by the very assumption that the quantum description is based on.

@@

As we went through many details in the derivation, it is time to step back and conclude with a few general remarks.

The first remarkable aspect of this work is that something along these lines can be done. Regardless of whether all the minor details are addressed in the best possible way, the idea that so much can be derived from so little was not something we expected. In hindsight, the idea that physics should be founded on physical principles is nothing new: it just hasn't been thoroughly applied to basic mechanics. If this work had been done incrementally from the start, these concepts would have been developed gradually and the results would feel more mundane.

The second striking aspect is a shift in perspective that this work seems to suggest. The usual reductionist view is that you start with the rules of the small parts and then build up the composite whole. This work flips that perspective: it's not what the parts can tell you about the whole, it's what the whole can tell you about the parts.

The state space for classical particles is the one that allows us to express invariant distributions: that is a requirement that comes from the whole. The fragments of quantum particles do not even have a well defined description: the whole tells us nothing about their configuration. In retrospect, this is not be surprising. Nobody is going to hand us down the laws of infinitesimal parts: we induce them by manipulating finite systems. Therefore what the whole allows us to infer about its part is physically fundamental.

@@

In this section we include mathematical demonstration in support of the physical justifications of the paper.

@@

\begin{thrm}\label{symplectomorphism_generator}
	Let $(M, \omega)$ a symplectic manifold. Let $f: (M, \omega) \rightarrow (M, \omega)$ an infinitesimal self-symplectomorphism. Let $S \in TM \; | \; f(\xi^a(m)) = \xi^a(m) + S^a(m)dt \; \forall m \in M$ be the infinitesimal displacement. There exists a function $H: M \rightarrow \mathbb{R}$ such that $S^{a} \omega_{ab} = \partial_{b}H$ and $H \in C^2(M, \mathbb{R})$.
\end{thrm}
